
# Middle-Square

The first PRNG algorithm we'll look at is one of the very first to be popularized (1949) and was credited to John von Neumann. The implementation for an n-digit number sequence is as follows

- Take a starting seed number and square it
- If the result has fewer than 2n digits, leading zeroes are added to compensate
- The middle n digits of the result become the next number

And that's it!

Here is an implementation in r:

```{r}
middle_square <- function(seed){
  n <- 4
  result <- toString(seed^2)
  while (nchar(result) < n * 2){
    result <- paste0("0", result)
  }
  return(as.numeric(substr(result, 3, 6)))
}
```

Lets take a look at it in R and see what kind of random sequences it produces. Our generator will take in a 4 digit number as the seed

```{r}
seed <- 2341
results <- seed
for (i in 1:1000) {
  seed <- middle_square(seed)
  results <- c(results, seed)
}
```



Histogram of our results to see the distribution of numbers
```{r, message=FALSE}
library(tidyverse)
ggplot() +
  geom_histogram(aes(results))
```

**OOF** what happened?

We started off good, but then hit a point which had a very small *peroid*. This means it repeats itself very quickly. As we can see from the histogram, we ran into a point which had a peroid of 4 and it kept looping between those numbers.  


